---
title: "Katja data visuals"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(ggalluvial)
library(tidyverse)
library(ggthemes)
```

```{r}
data <- read_xlsx("../data/Rawdata_Compositionality_18122019.xlsx", sheet = 1)
```
# Context to reaction

```{r}
data %>%
  group_by(#rel_dom,
           #face,
           context_dyad, 
           #GEST, 
           REACT)%>%
  summarise(freq = n())%>%
  filter(#rel_dom != "NA", 
         #face != "NA",
         context_dyad != "NA",
         #GEST != "NA",
         REACT != "NA")%>%
  ggplot(.,
       aes(axis1 = context_dyad, 
           #axis2 = face, 
           #axis3 = GEST, 
           axis4 = REACT,
           y = freq)) +
  scale_x_discrete(limits = c("Context","Reaction"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = context_dyad)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  scale_fill_colorblind()

```
Negative context leads almost exclusively to avoid reaction, but not the other way round.

# Context to reaction by gesture

```{r}
data %>%
  group_by(#rel_dom,
           #face,
           context_dyad, 
           GEST, 
           REACT)%>%
  summarise(freq = n())%>%
  filter(#rel_dom != "NA", 
         #face != "NA",
         context_dyad != "NA",
         #GEST != "NA",
         REACT != "NA")%>%
  ggplot(.,
       aes(axis1 = context_dyad, 
           #axis2 = face, 
           #axis3 = GEST, 
           axis4 = REACT,
           y = freq)) +
  scale_x_discrete(limits = c("Context","Reaction"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill =GEST)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal()+
  scale_fill_colorblind()

```

Gesture seems to be not very predictive when it comes to differentiating the path from positive to affiliate and positive to avoid. 

# Context to reaction by facial expression

```{r}
data %>%
  group_by(#rel_dom,
           face,
           context_dyad, 
           #GEST, 
           REACT)%>%
  summarise(freq = n())%>%
  filter(#rel_dom != "NA", 
         face != "NA",
         context_dyad != "NA",
         #GEST != "NA",
         REACT != "NA")%>%
  ggplot(.,
       aes(axis1 = context_dyad, 
           #axis2 = face, 
           #axis3 = GEST, 
           axis4 = REACT,
           y = freq)) +
  scale_x_discrete(limits = c("Context","Reaction"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill =face)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal()+
  scale_fill_colorblind()

```

Hoot facial expressions seem to be more pronounced in the positive to avoid trajectory


# Context to reaction by dominance

```{r}
data %>%
  group_by(rel_dom,
           #face,
           context_dyad, 
           #GEST, 
           REACT)%>%
  summarise(freq = n())%>%
  filter(rel_dom != "NA", 
         #face != "NA",
         context_dyad != "NA",
         #GEST != "NA",
         REACT != "NA")%>%
  ggplot(.,
       aes(axis1 = context_dyad, 
           #axis2 = face, 
           #axis3 = GEST, 
           axis4 = REACT,
           y = freq)) +
  scale_x_discrete(limits = c("Context","Reaction"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill =rel_dom)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal()+
  scale_fill_colorblind()

```

Senders who are *not* dominant get much more avoid reactions. In other words: Dominant individuals show more avoid reactions.

# Relation between context and dominance

```{r}
data %>%
  group_by(rel_dom,
           #face,
           #context_dyad, 
           #GEST, 
           REACT
           )%>%
  summarise(freq = n())%>%
  filter(rel_dom != "NA", 
         #face != "NA",
         #context_dyad != "NA",
         #GEST != "NA",
         REACT != "NA"
         )%>%
  ggplot(.,
       aes(#axis1 = context_dyad, 
           #axis2 = face, 
           #axis3 = GEST, 
           axis4 = REACT,
           axis5 = rel_dom,
           y = freq)) +
  scale_x_discrete(limits = c("Context","Dominance Relation"), expand = c(.2, .05)) +
  geom_alluvium() +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal()+
  scale_fill_colorblind()

```

